---
title: "Sunday"
author: "Lucy Eckert"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(caret)
library(ggplot2)
library(dplyr)
library(knitr)
library(party)
library(gbm)
# rmarkdown::render("Leckert_Proj2.Rmd", output_file="README.md")
```



```{r}
data.path <- "C:/Users/leckert/Documents/NCSU/ST558/Project_2"
day <- read_csv(paste0(data.path,"/day.csv"))
byday <- day %>% select(-c(casual,registered, instant, dteday))
#Filter out Monday data, and remove unused variables
ByDay <- day %>% filter(weekday==7) %>% select(-c(casual,registered, instant, dteday))
```

Review Data by Summaries and Plots
```{r, , warning = FALSE}
#Rentals by Temperature
a <- ggplot(ByDay, aes(temp, cnt))
a + geom_jitter() +geom_smooth() +labs(title = "Bike Rental Count by Temperature", 
                                       x = "Normalized Temperature in Celcius", 
                                       y = "Count of Bike Rentals")
```

```{r, , warning = FALSE}
#Rentals by Season
b <- ggplot(day, aes(x = season, y = cnt))
b + geom_bar(stat = "identity", aes(y=cnt, fill="Season"), colour="green") + labs(title = "Bike Rental Count by Season", x = "Season", y = "Count of Bike Rentals") + scale_fill_discrete(name = "Seasons:", 
       labels = c("Winter", "Spring", "Summer", "Fall")) 
```

```{r, warning = FALSE}
#Rentals by Weather Type
c <- ggplot(day, aes(x = weathersit, y = cnt))
c + geom_bar(stat = "identity", aes(y=cnt, fill="Weather"), colour="green") + 
  labs(title = "Bike Rental Count by Weather Type", x = "Weather Type", y = "Count of Bike Rentals") +   scale_fill_discrete(name = "Weather:") 
```

Review Summary Stats for Continuous Variables
```{r}
summary_data <- ByDay %>% select(temp:windspeed)
kable(apply(summary_data, 2, summary), caption = paste("Summary Stats for Continuous Variables"), 
      digit = 2)
```


Create train and test data sets for Monday data. Clean data.
```{r, , warning = FALSE}
set.seed(1)
trainIndex <- createDataPartition(ByDay$cnt, p = 0.7, list = FALSE)
Train <- ByDay[trainIndex, ] 
Test <-  ByDay[-trainIndex, ]
```
  
### Build Models forTrain Data  

**Model 1: Non-Ensemble Tree** 

While doing some research on model building, I discovered the concept of of using dummy 
variables as a way to create "switches" for some of the variables. It really helped me
break down which were more useful for the model.
```{r, warning = FALSE}
ByDay      <- day %>% filter(weekday==7) %>% select(-c(casual,registered, instant, dteday))
trainIndex   <- createDataPartition(ByDay$cnt, p = 0.7, list = FALSE)
Train <- ByDay[trainIndex, ] %>% select(-c(workingday, weekday)) %>% 
  mutate(mnth=as.factor(mnth), season=as.factor(season), weathersit = as.factor(weathersit))
dmy          <- dummyVars(" ~ .", data = Train, fullRank = T)
Train.trf <- data.frame(predict(dmy, newdata = Train)) %>% mutate(y = scale(cnt)) %>% select(-cnt)
fitControl <- trainControl(method = "LOOCV")
model      <- train(y ~., data = Train.trf, method = "ctree",
                    trControl = fitControl)
print(model)
plot(model$finalModel)
```

**Model 2: Boosted Tree**

I selected this model after trying many combinations of the n.trees, shrinkage, 
and interaction depth. I selected it for the most favorable RMSE.
```{r, warning = FALSE}
set.seed(1)
boostFit8 <- gbm(cnt ~., data = Train, distribution = "gaussian", n.trees = 100,
                 shrinkage = .1, interaction.depth = 2)
boostPred <- predict(boostFit8, newdata = dplyr::select(Test, -cnt), n.trees = 100)
boostRMSE <- sqrt(mean((boostPred-Test$cnt)^2))
#Print RMSE
boostRMSE