---
title: "Bike Data Project - ST558"
author: "Lucy Eckert"
date: "10/8/2020"
output:
  github_document:
    toc: true
    toc_depth: 4
---
# ST 558 Project #2, Group B - Predicting Bike Rental Totals  

## Introduction and Supporting Information  
### Introduction
    + I used data from the [Capital Bikeshare Program](https://www.capitalbikeshare.com/), 
    which is a bike sharing program based in the Metro D.C. area of the United States. 
    I reviewed data from 2011 and 2012, and used various predictors to create models 
    that would predict the daily total of bike rentals.  
    
### Supporting Information  
    + For this project, I am using a number of packages in R. They include, Tidyverse, 
    Readr, Caret, GGplot2, tree, dplyr, knitr, party, gbm,    
    
### Information about Variable Selection  

Below I have included a quick note about each variable. 
- **instant**: This variable has been removed from my prediction dataset as is simply a record index.  
- **dteday**: This variable has been removed from my prediction dataset as it does not contribute to the prediction.  

- **season**: Indicates season (1:winter, 2:spring, 3:summer, 4:fall). I have converted these to a dummy variable, to indicate the absence or presence of some categorical effect that may be expected to shift the outcome.  

- **yr**: Indicates year: (0: 2011, 1:2012)  

- **mnth**: Indicates month: ( 1 to 12)  

- **holiday**: Indicates whether day is holiday or not (extracted from [Web Link](https://dchr.dc.gov/page/holiday-schedules))  

- **weekday**: Day of the week, with Monday being 1.  

- workingday : Removed, as is complementary to holiday  

+ **weathersit**:  Indicates weather category. I have converted these to a dummy variable, to indicate the absence or presence of some categorical effect that may be expected to shift the outcome. Descriptions follow
- 1: Clear, Few clouds, Partly cloudy, Partly cloudy
- 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
- 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
- 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog  

- **temp** : Normalized temperature in Celsius.   

- **atemp**: Normalized feeling temperature in Celsius.   

- **hum**: Normalized humidity. The values are divided to 100 (max)  

- **windspeed**: Normalized wind speed. The values are divided to 67 (max)  

- **cnt**: The variable I am building models to predict, it is the count of total rental bikes including both casual and registered

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

library(tidyverse)
library(readr)
library(caret)
library(ggplot2)
library(dplyr)
library(knitr)
library(party)
library(gbm)
# rmarkdown::render("Leckert_Proj2.Rmd", output_file="README.md")
```

## Work with Data  

Create relative path, pull in data, and create Monday dataset.
```{r, , warning = FALSE}
data.path <- "C:/Users/leckert/Documents/NCSU/ST558/Project_2"
day <- read_csv(paste0(data.path,"/day.csv"))
#Filter out Monday data, and remove unused variables
Monday <- day %>% filter(weekday==1) %>% select(-c(casual,registered, instant, dteday))
```

Review Data by Summaries and Plots
```{r, , warning = FALSE}
#Rentals by Temperature
a <- ggplot(Monday, aes(temp, cnt))
a + geom_jitter() +geom_smooth() +labs(title = "Bike Rental Count by Temperature", 
                                       x = "Normalized Temperature in Celcius", 
                                       y = "Count of Bike Rentals")
```

```{r, , warning = FALSE}
#Rentals by Season
b <- ggplot(Monday, aes(x = season, y = cnt))
b + geom_bar(stat = "identity", aes(y=cnt, fill="Season"), colour="green") + labs(title = "Bike Rental Count by Season", x = "Season", y = "Count of Bike Rentals") + scale_fill_discrete(name = "Seasons:", 
       labels = c("Winter", "Spring", "Summer", "Fall")) 
```

```{r, warning = FALSE}
#Rentals by Weather Type
c <- ggplot(Monday, aes(x = weathersit, y = cnt))
c + geom_bar(stat = "identity", aes(y=cnt, fill="Weather"), colour="green") + 
  labs(title = "Bike Rental Count by Weather Type", x = "Weather Type", y = "Count of Bike Rentals") +   scale_fill_discrete(name = "Weather:") 
```

Review Summary Stats for Continuous Variables
```{r}
summary_data <- Monday %>% select(temp:windspeed)
kable(apply(summary_data, 2, summary), caption = paste("Summary Stats for Continuous Variables"), 
      digit = 2)
```


Create train and test data sets for Monday data. Clean data.
```{r, , warning = FALSE}
set.seed(1)
trainIndex <- createDataPartition(Monday$cnt, p = 0.7, list = FALSE)
Monday.Train <- Monday[trainIndex, ] 
Monday.Test <-  Monday[-trainIndex, ]
```
  
### Build Models for Monday Train Data  

Model 1: Non-Ensemble Tree
```{r, warning = FALSE}
Monday       <- day %>% filter(weekday==1) %>% select(-c(casual,registered, instant, dteday))
trainIndex   <- createDataPartition(Monday$cnt, p = 0.7, list = FALSE)
Monday.Train <- Monday[trainIndex, ] %>% select(-c(workingday, weekday)) %>% 
  mutate(mnth=as.factor(mnth), season=as.factor(season), weathersit = as.factor(weathersit))
dmy          <- dummyVars(" ~ .", data = Monday.Train, fullRank = T)
Monday.Train.trf <- data.frame(predict(dmy, newdata = Monday.Train)) %>% mutate(y = scale(cnt)) %>% select(-cnt)
fitControl <- trainControl(method = "LOOCV")
model      <- train(y ~., data = Monday.Train.trf, method = "ctree",
                    trControl = fitControl)
print(model)
plot(model$finalModel)
```

Model 2: Boosted Tree
ADD SELECTION DISCUSSION
```{r, warning = FALSE}
set.seed(1)
boostFit1 <- gbm(cnt ~., data = Monday.Train, distribution = "gaussian", n.trees = 100,
                 shrinkage = .1, interaction.depth = 2)
boostPred <- predict(boostFit1, newdata = dplyr::select(Monday.Test, -cnt), n.trees = 100)
boostRMSE <- sqrt(mean((boostPred-Monday.Test$cnt)^2))
#Print RMSE
boostRMSE
```

### Automate to Tuesday:Sunday Data
```{r}

```

