---
title: "Bike Data Project - ST558"
author: "Lucy Eckert"
date: "10/8/2020"
output:
  github_document:
    toc: true
    toc_depth: 4

---
# Creating a Model for Predicting Bike Sharing Numbers   

## Introduction and Supporting Information  

### Introduction
    + I used data from the [Capital Bikeshare Program](https://www.capitalbikeshare.com/), 
    which is a bike sharing program based in the Metro D.C. area of the United States. 
    I reviewed data from 2011 and 2012, and used various predictors to create models 
    that would predict the daily total of bike rentals.  
    
    
### Supporting Information  
    + For this project, I am using a number of packages in R. They include, Tidyverse, 
    Readr, Caret, GGplot2, tree, dplyr, knitr, party, gbm,    
    
### Information about Variable Selection  

Below I have included a quick note about each variable. 
- **instant**: This variable has been removed from my prediction dataset as is simply a record index.  
- **dteday**: This variable has been removed from my prediction dataset as it does not contribute to the prediction.  

- **season**: Indicates season (1:winter, 2:spring, 3:summer, 4:fall). I have converted these to a dummy variable, to indicate the absence or presence of some categorical effect that may be expected to shift the outcome.  

- **yr**: Indicates year: (0: 2011, 1:2012)  

- **mnth**: Indicates month: ( 1 to 12)  

- **holiday**: Indicates whether day is holiday or not (extracted from [Web Link](https://dchr.dc.gov/page/holiday-schedules))  

- **weekday**: Day of the week, with Monday being 1.  

- workingday : Removed, as is complementary to holiday  

+ **weathersit**:  Indicates weather category. I have converted these to a dummy variable, to indicate the absence or presence of some categorical effect that may be expected to shift the outcome. Descriptions follow
- 1: Clear, Few clouds, Partly cloudy, Partly cloudy
- 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
- 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
- 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog  

- **temp** : Normalized temperature in Celsius.   

- **atemp**: Normalized feeling temperature in Celsius.   

- **hum**: Normalized humidity. The values are divided to 100 (max)  

- **windspeed**: Normalized wind speed. The values are divided to 67 (max)  

- **cnt**: The variable I am building models to predict, it is the count of total rental bikes including both casual and registered

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

library(tidyverse)
library(readr)
library(caret)
library(ggplot2)
library(dplyr)
library(knitr)
library(party)
library(gbm)
# rmarkdown::render("Leckert_Proj2.Rmd", output_file="README.md")
```

## Work with Data  

Create relative path, pull in data, and create Monday dataset.
```{r, , warning = FALSE}
day <- read_csv(file = "day.csv")
byday <- day %>% select(-c(casual,registered, instant, dteday))
#Filter out Monday data, and remove unused variables
Sunday <- day %>% filter(weekday==7) %>% select(-c(casual,registered, instant, dteday))
```

Review Data by Summaries and Plots
```{r, , warning = FALSE}
#Rentals by Temperature
a <- ggplot(Sunday, aes(temp, cnt))
a + geom_jitter() +geom_smooth() +labs(title = "Bike Rental Count by Temperature", 
                                       x = "Normalized Temperature in Celcius", 
                                       y = "Count of Bike Rentals")
```

```{r, , warning = FALSE}
#Rentals by Season
b <- ggplot(Sunday, aes(x = season, y = cnt))
b + geom_bar(stat = "identity", aes(y=cnt, fill="Season"), colour="green") + labs(title = "Bike Rental Count by Season", x = "Season", y = "Count of Bike Rentals") + scale_fill_discrete(name = "Seasons:", 
       labels = c("Winter", "Spring", "Summer", "Fall")) 
```

```{r, warning = FALSE}
#Rentals by Weather Type
c <- ggplot(Sunday, aes(x = weathersit, y = cnt))
c + geom_bar(stat = "identity", aes(y=cnt, fill="Weather"), colour="green") + 
  labs(title = "Bike Rental Count by Weather Type", x = "Weather Type", y = "Count of Bike Rentals") +   scale_fill_discrete(name = "Weather:") 
```

Review Summary Stats for Continuous Variables
```{r}
summary_data <- Sunday %>% select(temp:windspeed)
kable(apply(summary_data, 2, summary), caption = paste("Summary Stats for Continuous Variables"), 
      digit = 2)
```


Create train and test data sets for Monday data. Clean data.
```{r, , warning = FALSE, error=TRUE}
set.seed(1)
trainIndex <- createDataPartition(Sunday$cnt, p = 0.7, list = FALSE)
Sunday.Train <- Sunday[trainIndex, ] 
Sunday.Test <-  Sunday[-trainIndex, ]
```
  
### Build Models for Monday Train Data  

**Model 1: Non-Ensemble Tree** 

While doing some research on model building, I discovered the concept of of using dummy 
variables as a way to create "switches" for some of the variables. It really helped me
break down which were more useful for the model.
```{r, warning = FALSE, error=TRUE}
Sunday       <- day %>% filter(weekday==6) %>% select(-c(casual,registered, instant, dteday))
trainIndex   <- createDataPartition(Sunday$cnt, p = 0.7, list = FALSE)
Sunday.Train <- Sunday[trainIndex, ] %>% select(-c(workingday, weekday)) %>% 
  mutate(mnth=as.factor(mnth), season=as.factor(season), weathersit = as.factor(weathersit))
dmy          <- dummyVars(" ~ .", data = Sunday.Train, fullRank = T)
Sunday.Train.trf <- data.frame(predict(dmy, newdata = Sunday.Train)) %>% mutate(y = scale(cnt)) %>% select(-cnt)
fitControl <- trainControl(method = "LOOCV")
model      <- train(y ~., data = Sunday.Train.trf, method = "ctree",
                    trControl = fitControl)
print(model)
plot(model$finalModel)
```

**Model 2: Boosted Tree**

I selected this model after trying many combinations of the n.trees, shrinkage, 
and interaction depth. I selected it for the most favorable RMSE.
```{r, warning = FALSE, error=TRUE}
set.seed(1)
boostFit8 <- gbm(cnt ~., data = Sunday.Train, distribution = "gaussian", n.trees = 100,
                 shrinkage = .1, interaction.depth = 2)
boostPred <- predict(boostFit8, newdata = dplyr::select(Sunday.Test, -cnt), n.trees = 100)
boostRMSE <- sqrt(mean((boostPred-Sunday.Test$cnt)^2))
#Print RMSE
boostRMSE
```

**Model 3: Linear Regression**
```{r regmodel, echo=TRUE, warning=FALSE, error=TRUE}
set.seed(1)
regFit <- lm(cnt ~ temp + hum + windspeed, data=Sunday.Train)

regPred <- predict(regFit, newdata = dplyr::select(Sunday.Test, -cnt))
```


### Code Automation  

I'm running this code in another rmd to automate the other reports, but I wanted to include
it here (with eval = FALSE) so that my project is documented in one place.  

```{r, eval = FALSE}
dayofweek <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

output_file <- paste0(dayofweek, "Analysis", ".md")

params <- lapply(dayofweek, FUN = function(x){list(weekday = x)})

reports <- tibble(output_file, params)

apply(reports, MARGIN = 1, FUN = function(x){render(input = "Leckert_Proj2.Rmd", output_file = x[[1]], params = x[[2]])})
```

The analysis for [Monday can be found here.](https://github.com/laeckert/Project_2/blob/main/MondayAnalysis_files/figure-gfm/Monday.pdf)

The analysis for [Tuesday can be found here.](https://github.com/laeckert/Project_2/blob/main/TuesdayAnalysis_files/figure-gfm/Tuesday.pdf)

The analysis for [Wednesday can be found here.](https://github.com/laeckert/Project_2/blob/main/WednesdayAnalysis_files/figure-gfm/Wednesday.pdf)

The analysis for [Thursday can be found here.](https://github.com/laeckert/Project_2/blob/main/ThursdayAnalysis_files/figure-gfm/Thursday.pdf)

The analysis for [Friday can be found here.](https://github.com/laeckert/Project_2/blob/main/FridayAnalysis_files/figure-gfm/Friday.pdf)

The analysis for [Saturday can be found here.](https://github.com/laeckert/Project_2/blob/main/SaturdayAnalysis_files/figure-gfm/Saturday.pdf)

The analysis for [Sunday can be found here.](https://github.com/laeckert/Project_2/blob/main/SundayAnalysis_files/figure-gfm/Sunday.pdf)
